---
title: "Final Project Report"
subtitle: "DATA2020 Statistical Learning"
author:
  - Jiaxin Guan
  - Qiuli Lai
  - Yitong Liu
date: "April 2026"
output:
  pdf_document:
    latex_engine: xelatex
fontsize: 11pt
geometry: margin=1in
---

\begin{center}
Brown University \\
Data Science Initiative
\end{center}

\vspace{0.5cm}

\textbf{Instructor}: Alice Paul

\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  fig.width = 7,
  fig.height = 4.5
)
```

```{r libraries}
library(tidyverse)
library(caret)
library(ggplot2)
library(patchwork)
```

------------------------------------------------------------------------

# 1. Introduction and Background

## 1.1 Problem Context

-   **The problem addressed in this paper is:** Predicting the financial success (first-weekend box office revenue) of movies *before* their release date, using only passively collected online activity data, specifically, Wikipedia editing and viewing behavior.

-   **Why this problem matters:** The movie industry involves enormous financial stakes. Studios and distributors make multimillion-dollar decisions about marketing spend, theater allocation, and release timing based on uncertain forecasts. Traditional forecasting relies on expert judgment or pre-release surveys, which are costly and slow. A data-driven early prediction system could provide actionable signals weeks before release.

-   **Real-world implication:** Marketing departments, distributors, and theater chains could benefit from more accurate pre-release revenue forecasts. The methodology is also generalizable to other cultural products (books, music, games) and to non-English-speaking markets.

------------------------------------------------------------------------

## 1.2 Paper Overview

-   **The paper studies:** The relationship between Wikipedia activity (page views, number of editors, number of edits, collaborative rigor) for a movie's Wikipedia article and its eventual first-weekend box office revenue in the United States.

-   **The main idea is:** Collective online behavior on Wikipedia which is a platform where engaged users proactively write and edit movie articles months before release, serves as a leading indicator of public interest and financial success.

------------------------------------------------------------------------

## 1.3 Research Question

-   **The research question is:** Can Wikipedia activity data be used to predict a movie's first-weekend box office revenue, and if so, how early before release can a reasonable prediction be made?

------------------------------------------------------------------------

## 1.4 Hypothesis

-   **The authors hypothesize that:** Higher levels of Wikipedia editing and viewing activity for a movie's article, measured weeks to months before release, will be positively correlated with higher first-weekend box office revenue, and that a multivariate linear regression model using these activity measures will outperform a benchmark model based solely on the number of theaters screening the film.

------------------------------------------------------------------------

# 2. Methods

## 2.1 Replication Methods

### 2.1.1 Data Source

-   **Data comes from:** Two sources: (1) Box Office Mojo, for first-weekend box office revenue and theater counts for U.S. releases in 2010; (2) Wikimedia Toolserver and Wikipedia article traffic statistics, for per-day cumulative editing and viewing activity on each movie's Wikipedia page.

-   **Unit of observation:** One movie (one row per film, with associated time-series predictor files).

-   **Sample size:** 312 movies (out of 535 total U.S. releases in 2010 that could be matched to a Wikipedia page). We also work with a sub-sample of 24 movies (the Asur & Huberman Twitter study sample) for direct comparison.

------------------------------------------------------------------------

### 2.1.2 Variables

**Response Variable**

-   $y$: First-weekend box office revenue in the U.S. (USD)

**Predictors** (all measured as cumulative values from article inception up to prediction time $t$)

-   $V$: Number of page views
-   $U$: Number of distinct human editors (users)
-   $E$: Number of edits by human editors
-   $R$: Collaborative rigor — like $E$ but counts consecutive edits by the same user as one (reduces noise from single-user editing bursts)
-   $T$: Number of theaters screening the movie on opening weekend (benchmark variable; not Wikipedia-derived)

------------------------------------------------------------------------

### 2.1.3 Data Preprocessing

-   **Step 1:** Load the movie index file (`sample_of_312`) containing revenue, theater count, release date, and Wikipedia article title.
-   **Step 2:** For each movie, load its individual Wikipedia predictor file from `wikipedia_predictors/`. Each file contains daily cumulative values of $V$, $U$, $R$, $E$ from $t = -500$ days to post-release.
-   **Step 3:** For each prediction time point $t$ (days before release, in "movie time"), extract the predictor values at that $t$ for all 312 movies, constructing a cross-sectional dataset.

------------------------------------------------------------------------

### 2.1.4 Statistical Model

-   **Model used:** Multivariate linear regression

-   **Model specification:** $$y = \sum_{j \in S} \alpha_j(t)\, x_j(t) + C_S(t) + \varepsilon_S(t)$$ where $S$ is the set of predictor variables, $\alpha_j(t)$ are time-varying regression coefficients, $C_S(t)$ is the intercept, and $\varepsilon_S(t)$ is the error term.

-   **Reason:** Linear regression is interpretable, fast to compute across hundreds of time points, and sufficient to demonstrate that Wikipedia activity carries predictive signal. The authors explicitly favor a "minimalistic" approach.

-   **Replication consistency:** We replicate using the same predictor variable definitions, the same sample, and the same evaluation scheme (10-fold cross-validation for $R^2$).

------------------------------------------------------------------------

```{r load-data}
index_path   <- "data/wikipredict_data_pack/sample_of_312/sample_of_312"
pred_dir_312 <- "data/wikipredict_data_pack/sample_of_312/wikipedia_predictors/"
index_path24 <- "data/wikipredict_data_pack/asur_huberman_sample_of_24/asur_huberman_sample_of_24"
pred_dir_24  <- "data/wikipredict_data_pack/asur_huberman_sample_of_24/wikipedia_predictors/"

# load movie index
index312 <- read.table(index_path, sep = "\t", header = TRUE, quote = "")
colnames(index312) <- c("ID", "Title", "WP_title", "Revenue", "Theaters",
                        "Release_date", "Inception_days")

cat("Sample size:", nrow(index312), "\n")
cat("Revenue range: $", min(index312$Revenue), "–$", max(index312$Revenue), "\n")
head(index312[, c("ID", "Title", "Revenue", "Theaters")], 5)
```

```{r load-predictors}
# load all predictor time-series into memory
load_all_predictors <- function(id_vec, pred_dir) {
  preds <- list()
  for (id in id_vec) {
    fp <- file.path(pred_dir, as.character(id))
    if (file.exists(fp)) {
      d <- tryCatch(
        read.table(fp, sep = "\t", header = TRUE, quote = ""),
        error = function(e) NULL
      )
      if (!is.null(d)) preds[[as.character(id)]] <- d
    }
  }
  preds
}

preds312 <- load_all_predictors(index312$ID, pred_dir_312)
cat("Loaded predictor files:", length(preds312), "/ 312\n")
```

------------------------------------------------------------------------

## 2.2 Extension Methods

### 2.2.1 Motivation

-   **Limitation of original paper:** The original model applies linear regression directly to raw (untransformed) predictor values. However, as visible from the log-log axes in Figure 5 of the paper, the relationship between Wikipedia activity and box office revenue appears log-linear (spanning many orders of magnitude). Using raw values violates the homoscedasticity assumption of OLS regression and may produce poor predictions for low-revenue films.

-   **Proposed extension:**

------------------------------------------------------------------------

### 2.2.2 New Features

### 2.2.3 Model

### 2.2.4 Justification

# 3. Results

## 3.1 Replication Results

```{r figure1, fig.cap="Figure 1 Replication: Histograms of different variables for the sample of n=312 movies from 2010. A: time of Wikipedia article creation (movie time days); B: first-weekend box office revenue (USD); C: number of theaters; D–G: accumulated Wikipedia activity up to t=7 days after release.", fig.width=9, fig.height=5}
library(patchwork)
 
# ── Get predictor values accumulated up to t = +7 days after release ──────────
df7 <- get_at_time(7, preds312, index312)
 
# ── Shared histogram theme (matching original paper: white bars, black border) ─
ht <- theme_classic(base_size = 9) +
  theme(
    axis.title.x    = element_text(size = 9),
    axis.title.y    = element_text(size = 8),
    axis.text       = element_text(size = 7),
    plot.tag        = element_text(face = "bold", size = 10),
    plot.tag.position = "topleft"
  )
bar_fill  <- "grey"
bar_color <- "black"
bar_lwd   <- 0.3
 
# ── Panel A: Article inception time (movie time, days) ────────────────────────
pA <- ggplot(data.frame(tc = index312$Inception_days), aes(x = tc)) +
  geom_histogram(bins = 35, fill = bar_fill, color = bar_color,
                 linewidth = bar_lwd) +
  labs(x = expression(t[c]), y = "Count of movies", tag = "A") +
  ht
 
# ── Panel B: Box office revenue (log10) ───────────────────────────────────────
pB <- ggplot(df7, aes(x = log10(Revenue))) +
  geom_histogram(bins = 20, fill = bar_fill, color = bar_color,
                 linewidth = bar_lwd) +
  labs(x = expression(log[10](Y)), y = "", tag = "B") +
  ht
 
# ── Panel C: Number of theaters (log10) ───────────────────────────────────────
pC <- ggplot(df7, aes(x = log10(Theaters + 1))) +
  geom_histogram(bins = 15, fill = bar_fill, color = bar_color,
                 linewidth = bar_lwd) +
  scale_x_continuous(breaks = 0:3) +
  labs(x = expression(log[10](T)), y = "", tag = "C") +
  ht
 
# ── Panel D: Accumulated views (log10) ────────────────────────────────────────
pD <- ggplot(df7, aes(x = log10(V + 1))) +
  geom_histogram(bins = 20, fill = bar_fill, color = bar_color,
                 linewidth = bar_lwd) +
  labs(x = expression(log[10](V)), y = "Count of movies", tag = "D") +
  ht
 
# ── Panel E: Accumulated users (log10) ────────────────────────────────────────
pE <- ggplot(df7, aes(x = log10(U + 1))) +
  geom_histogram(bins = 20, fill = bar_fill, color = bar_color,
                 linewidth = bar_lwd) +
  scale_x_continuous(breaks = 0:3) +
  labs(x = expression(log[10](U)), y = "", tag = "E") +
  ht
 
# ── Panel F: Accumulated edits (log10) ────────────────────────────────────────
pF <- ggplot(df7, aes(x = log10(E + 1))) +
  geom_histogram(bins = 20, fill = bar_fill, color = bar_color,
                 linewidth = bar_lwd) +
  scale_x_continuous(breaks = 0:3) +
  labs(x = expression(log[10](E)), y = "", tag = "F") +
  ht
 
# ── Panel G: Accumulated rigor (log10) ────────────────────────────────────────
pG <- ggplot(df7, aes(x = log10(R + 1))) +
  geom_histogram(bins = 20, fill = bar_fill, color = bar_color,
                 linewidth = bar_lwd) +
  scale_x_continuous(breaks = 0:3) +
  labs(x = expression(log[10](R)), y = "", tag = "G") +
  ht
 
# ── Combine: top row A B C, bottom row D E F G ────────────────────────────────
(pA | pB | pC) / (pD | pE | pF | pG)
```

```{r pearson-correlation, fig.cap="Figure 2 Replication: Pearson correlation of box office revenue with each predictor over time."}
time_points <- seq(-400, 100, by = 2)

cor_results <- bind_rows(lapply(time_points, function(t) {
  df <- get_at_time(t, preds312, index312)
  data.frame(
    t = t,
    V        = cor(df$Revenue, df$V,        use = "complete.obs"),
    U        = cor(df$Revenue, df$U,        use = "complete.obs"),
    R        = cor(df$Revenue, df$R,        use = "complete.obs"),
    E        = cor(df$Revenue, df$E,        use = "complete.obs"),
    Theaters = cor(df$Revenue, df$Theaters, use = "complete.obs")
  )
}))

# reshape
cor_long <- reshape(
  cor_results,
  varying   = c("V", "U", "R", "E", "Theaters"),
  v.names   = "r",
  timevar   = "Variable",
  times     = c("V", "U", "R", "E", "Theaters"),
  direction = "long"
)

cor_long$Variable <- factor(
  cor_long$Variable,
  levels = c("V", "U", "R", "E", "Theaters")
)

# style
var_colors <- c(
  V = "#2C3E90",
  U = "#C0392B",
  R = "#E67E22",
  E = "#F1C40F",
  Theaters = "#636363"
)

var_lty <- c(
  V = "solid",
  U = "solid",
  R = "solid",
  E = "solid",
  Theaters = "dashed"
)

var_labels <- c(
  V = "V",
  U = "U",
  R = "R",
  E = "E",
  Theaters = "T"
)


main_p <- ggplot(cor_long, aes(x = t, y = r,
                               color = Variable, linetype = Variable)) +
  geom_line(linewidth = 0.5) +
  scale_color_manual(values = var_colors, labels = var_labels) +
  scale_linetype_manual(values = var_lty, labels = var_labels) +
  scale_x_continuous(
    limits = c(-400, 100),
    breaks = c(-400, -300, -200, -100, 0, 100),
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    limits = c(0.5, 1.001),
    breaks = seq(0.5, 1.0, by = 0.1),
    expand = c(0, 0)
  ) +
  labs(
    x = "Time (day)",
    y = expression(r[j](t)),
    color = NULL,
    linetype = NULL
  ) +
  theme_classic(base_size = 10) +
  theme(
    legend.position   = c(0.93, 0.18),
    legend.text.align = 1,
    legend.key.width  = unit(1.2, "cm"),
    legend.key.height = unit(0.4, "cm"),
    legend.spacing.y  = unit(0.1, "cm"),
    panel.border      = element_rect(color = "black", fill = NA, linewidth = 0.5),
    axis.line         = element_blank()
  ) +
  guides(
    color    = guide_legend(label.position = "left"),
    linetype = guide_legend(label.position = "left")
  )


inset_data <- cor_long[cor_long$t >= -20 & cor_long$t <= 20, ]


inset_p <- ggplot(inset_data, aes(x = t, y = r,
                                  color = Variable, linetype = Variable)) +
  geom_line(linewidth = 0.4) +
  scale_color_manual(values = var_colors) +
  scale_linetype_manual(values = var_lty) +
  scale_x_continuous(
    limits = c(-20, 20),                
    breaks = c(-20, -10, 0, 10, 20),
    expand = c(0, 0)                     
  ) +
  scale_y_continuous(
    limits = c(0.74, 0.86),             
    breaks = c(0.75, 0.80, 0.85),
    expand = c(0, 0)
  ) +
  theme_classic(base_size = 8) +
  theme(
    legend.position = "none",
    panel.border    = element_rect(color = "black", fill = NA, linewidth = 0.4),
    axis.line       = element_blank(),
    plot.background = element_rect(fill = "white", color = "black", linewidth = 0.4)
  )

main_p +
  annotation_custom(
    grob = ggplotGrob(inset_p),
    xmin = -400,
    xmax = -200,  
    ymin = 0.78,
    ymax = 1
  )
```

## 3.2 Extension Results

------------------------------------------------------------------------

# 4. Discussion and Conclusion

## 4.1 Interpretation

------------------------------------------------------------------------

## 4.2 Strengths

------------------------------------------------------------------------

## 4.3 Limitations

## 4.4 Challenges

## 4.5 Future Work

## 4.6 Final Takeaway
